---
title: "Predict activity quality from activity monitors through motion data "
author: "Aamarcha"
date: "23 août 2014"
---
The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. The goal is to predict the activity quality from activity monitors collected from these exercices. Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes.
Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3BIcn6MnR

The accuracy of the prediction of the activity quality is arount 99.5% using Random forest algorithm.

#Methodoloy

1. First I clean data : removing NA and not generalized variables, classes as factors.
2. Then I make a partition : 75% for traing data and 25% for test.
3. I build a model using random forest algorithm. GMB implementation takes a very long time.
4. I make predictions basing on this Model.

#Getting Data
Data was downloaded from 

1. Training data : https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
2. Test data : https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The package used for prediction is caret.
```{r loadPackages, results="hide",message=FALSE}
library(caret)
library(plyr)
library(randomForest)
library('scales')
```

Reading the data:
```{r preparing, cache=TRUE, results="hide"}
trainingMotion <- read.csv("./data/pml-training.csv",stringsAsFactors=F,na.string = c("NA", ""))
testMotion <- read.csv("./data/pml-testing.csv",stringsAsFactors=F,na.string = c("NA", ""))
```
#Cleaning and preparing data
Remove NA data and not generalizable colomns. As you can see, 160 variables of the data contains 98% missing or NA value.
```{r cleanNA,}
NAcounter  <- apply(trainingMotion, 2, function(x) sum(is.na(x)))
table(round(NAcounter/nrow(trainingMotion), 2))
trainingMotion  <- trainingMotion[, names(NAcounter)[NAcounter == 0]]
trainingMotion <- trainingMotion[, -c(1:7)]
```

Use classes as a factor.
```{r cleanno,}
trainingMotion <- mutate(trainingMotion,classe=factor(classe))
```
Create partitions, 75% for traning data and 2% for test data.

```{r partition,}
set.seed(123)
inTrain <- createDataPartition(y = trainingMotion$classe, p = 0.75, list = F)
trainingMotion.train <- trainingMotion[inTrain, ]
trainingMotion.test <- trainingMotion[-inTrain, ]
```

# Model building
I use the random forest implementation because it's faster in my context. You can see that the model is 99.5% accurate and  the 95% confidence interval for accuracy is from 99.3% to 99.7%.

```{r build,}
set.seed(456)
# repeated cross validation
trControl <- trainControl(method="repeatedcv", number=10, repeats=10)

# train with  "rf" method takes a long time
rf.mdl<- randomForest(classe ~ ., data = trainingMotion.train,verbose=FALSE,metric="Accuracy", trControl =trControl )
confusionMatrix <- confusionMatrix(trainingMotion.test$classe, predict(rf.mdl, trainingMotion.test))
confusionMatrix
```

Here you have the error plot and the importance of variables in this model.

```{r plot,}
plot(rf.mdl, main="Error vs trees")
varImpPlot(rf.mdl, main="Importance of variables")
```

Out of sample error rate
```{r error,}
percent(1 - confusionMatrix$overall[[1]])
```

#Prediction and submission
```{r prediction,}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
prediction  <- predict(rf.mdl, testMotion)
prediction
pml_write_files(prediction)
```

#Appendix
The importance of variables in this model
```{r imprtance,}
importance(rf.mdl)
```

